The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging with the Owens Valley Long Wavelength Array

### 2017-06-26 paper.tex
fix some unicode errors and minor changes

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}

@@

The Sun can be trivially removed from any map of the sky by constructing the map using only data
taken at night. A map of the entire sky can be obtained by using observations spaced 6 months apart.
However the OVRO-LWA 100 hour dataset consists of 100 consecutive hours of data, and hence roughly
half of the data is from the day.

The Sun is a time variable source in radio.

\todo{does the previous routine work on the Sun too? worth a shot}

@@

\begin{figure*}[ht]
    \centering
    \begin{tabular}{cccc}
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw04-small} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw06-small} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw08-small} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw10-small} \\
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw12-small} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw14-small} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw16-small} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw18-small} \\
    \end{tabular}
    \caption{
        Dirty snapshot images after smearing the visiblities over the entire 100 hour observation
        window.  The two bright source tracks correspond to Cas A (top track) and Cyg A (bottom
        track) smeared along a ring of constant declination. The sources dotting the horizon are
        faulty power line insulators.
    }
\end{figure*}

@@

Naively one might attempt to solve equation~\ref{eq:basic-imaging} by discretizing, and subsequently
solving the resulting matrix equation. If the interferometer is composed of $N_\text{base}$
baselines, and measures $N_\text{freq}$ frequency channels over $N_\text{time}$ integrations then
the entire data set consists of $N_\text{base}N_\text{freq}N_\text{time}$ complex numbers. If the
sky is discretized into $N_\text{pix}$ pixels then the relevant matrix has dimensions of
$(N_\text{base}N_\text{freq}N_\text{time})\times(N_\text{pix})$. For making single-channel maps with
the Owens Valley LWA this becomes a 5 petabyte array (assuming each matrix element is a 64-bit
complex floating point number).  This matrix equation is therefore prohibitively large, and solving
equation~\ref{eq:basic-imaging} by means of discretization is usually impossible although
\citet{2017MNRAS.465.2901Z} demonstrate this technique with the MITEOR telescope.

Instead it is common to make mild assumptions that simplify equation~\ref{eq:basic-imaging} and ease
the computational burden in solving for $I_\nu(\hat r)$. For example, when all of the baselines
$\vec b_{ij}$ lie in a plane and the field-of-view is small, equation~\ref{eq:basic-imaging} can be
well-approximated by a two-dimensional Fourier transform \citep{2001isra.book.....T}.

\todo{discuss w-projection}

\todo{discuss a-projection}

@@

In this context we will define a transit telescope as any interferometer where the response pattern
of the individual elements does not change with respect to time. This may be an interferometer like
the OVRO-LWA where the correlation elements are fixed dipoles, but it may also be an interferometer
like the MWA if the steerable beams are held in a fixed position (not necessarily at zenith).

@@

The interferometer's sensitivity to $l=0$, however, deserves special consideration.
\citet{2016ApJ...826..116V} prove -- under fairly general assumptions -- that an interferometer is
only sensitive to the monopole of the sky brightness if there exists some form of cross-talk or
common-mode noise. In fact, the sensitivity of the interferometer is proportional to the amplitude
of these effects. For consistency we will include $a_{00}$ while solving
Equation~\ref{eq:m-mode-matrix-equation} for the vector $\b a$, but we set $a_{00} = 0$ afterwards
because we do not have a measurement of cross-talk or common-mode noise.

@@

Imaging in $m$-mode analysis essentially amounts to inverting
equation~\ref{eq:m-mode-matrix-equation} to solve for the spherical harmonic coefficients $\b a$.
The linear-least squares solution, which minimizes $\|\b v - \b B\b a\|^2$, is given by
\begin{equation}
    \b{\hat a}_\text{LLS} = (\b B^*\b B)^{-1}\b B^*\b v\,,
\end{equation}
where $^*$ indicates the conjugate-transpose. However, usually one will find that $\b B$ is not
full-rank and hence $\b B^*\b B$ is not an invertible matrix. For example, an interferometer located
in the northern hemisphere will never see a region of the southern sky centered on the southern
celestial pole. The $m$-modes contained in the vector $\b v$ must contain no information
about the sky around the southern celestial pole, and therefore the act of multiplying by $\b B$
must destroy some information about the sky. The consequence of this fact is that $\b B^*\b B$ must
have at least one zero eigenvalue, which means it is not invertible.

@@

\subsection{$m$-Mode Analysis with the OVRO-LWA}

\subsection{CLEAN}

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}

@@

The Sun can be trivially removed from any map of the sky by constructing the map using only data
taken at night. A map of the entire sky can be obtained by using observations spaced 6 months apart.
However the dataset used in this paper consists of 28 consecutive hours. Additionally while most
astronomical sources at these frequencies have negative spectral indices, the Sun has a large
positive spectral index. Therefore more care will need to be taken in subtracting the Sun at higher
frequencies than at lower frequencies.

\todo{describe sun removal -- this isn't entirely settled yet}

@@

%\begin{figure*}[ht]
%    \centering
%    \begin{tabular}{cccc}
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw04-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw06-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw08-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw10-small} \\
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw12-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw14-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw16-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw18-small} \\
%    \end{tabular}
%    \caption{
%        Dirty snapshot images after smearing the visiblities over the entire 100 hour observation
%        window.  The two bright source tracks correspond to Cas A (top track) and Cyg A (bottom
%        track) smeared along a ring of constant declination. The sources dotting the horizon are
%        faulty power line insulators.
%    }
%\end{figure*}

@@

Naively one might attempt to solve equation~\ref{eq:basic-imaging} by discretizing, and subsequently
solving the resulting matrix equation. If the interferometer is composed of $N_\text{base}$
baselines, and measures $N_\text{freq}$ frequency channels over $N_\text{time}$ integrations then
the entire data set consists of $N_\text{base}N_\text{freq}N_\text{time}$ complex numbers. If the
sky is discretized into $N_\text{pix}$ pixels then the relevant matrix has dimensions of
$(N_\text{base}N_\text{freq}N_\text{time})\times(N_\text{pix})$. For making single-channel maps with
the OVRO-LWA this becomes a 5 petabyte array (assuming each matrix element is a 64-bit complex
floating point number).  This matrix equation is therefore prohibitively large, and solving
equation~\ref{eq:basic-imaging} by means of discretization is usually impossible although
\citet{2017MNRAS.465.2901Z} demonstrate this technique with the MITEOR telescope.

Instead it is common to make mild assumptions that simplify equation~\ref{eq:basic-imaging} and ease
the computational burden in solving for $I_\nu(\hat r)$. For example, when all of the baselines
$\vec b_{ij}$ lie in a plane and the field-of-view is small, equation~\ref{eq:basic-imaging} can be
well-approximated by a two-dimensional Fourier transform \citep{2001isra.book.....T}. The
restriction on baseline co-planarity and field-of-view can be relaxed by using W-projection
\citep{2008ISTSP...2..647C}. Known primary beam effects can also be accounted for during imaging by
using A-projection \citep{2013ApJ...770...91B}.

@@

In this context we will define a transit telescope as any interferometer where the response pattern
of the individual elements does not change with respect to time. This may be an interferometer like
the OVRO-LWA where the correlation elements are fixed dipoles, but it may also be an interferometer
like LOFAR or the MWA if the steerable beams are held in a fixed position (not necessarily at
zenith).

@@

The interferometer's sensitivity to $l=0$, however, deserves special consideration.
\citet{2016ApJ...826..116V} prove -- under fairly general assumptions -- that an interferometer is
only sensitive to the monopole of the sky brightness if there exists some form of cross-talk or
common-mode noise. In fact, the sensitivity of the interferometer is proportional to the amplitude
of these effects. For consistency we will include $a_{00}$ while solving
Equation~\ref{eq:m-mode-matrix-equation} for the vector $\b a$, but we set $a_{00} = 0$ afterwards
because we do not have a measurement of the amplitude of cross-talk or common-mode noise, which
would be needed to correctly calibrate $a_{00}$.

@@

Imaging in $m$-mode analysis essentially amounts to inverting
equation~\ref{eq:m-mode-matrix-equation} to solve for the spherical harmonic coefficients $\b a$.
The linear-least squares solution, which minimizes $\|\b v - \b B\b a\|^2$, is given by
\begin{equation}
    \b{\hat a}_\text{LLS} = (\b B^*\b B)^{-1}\b B^*\b v\,,
\end{equation}
where $^*$ indicates the conjugate-transpose. However, usually one will find that $\b B$ is not
full-rank and hence $\b B^*\b B$ is not an invertible matrix. For example, an interferometer located
in the northern hemisphere will never see a region of the southern sky centered on the southern
celestial pole. The $m$-modes contained in the vector $\b v$ must contain no information
about the sky around the southern celestial pole, and therefore the act of multiplying by $\b B$
must destroy some information about the sky. The consequence of this fact is that $\b B$ must have
at least one singular value that is equal to zero. It is then trivial to show that $\b B^*\b B$ must
have at least one eigenvalue that is equal to zero, which means it is not an invertible matrix.

@@

\subsection{Image Deconvolution}
